Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-19 article.tex
re-order text

\section{Experiments}
\label{sec:results}

@@

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{images/curves}
\caption{Method precision-recall and roc curves.}
\label{fig:curves}
\end{figure}

In this section, we will first describe datasets and methods used for the
comparison. Then, on Tab.\ref{tab:tab3}, our approach is decomposed to show
the contribution of each stages . Afterwards, we will perform a comparison
between several methods on Fig.\ref{fig:curves} and Tab.\ref{tab:tab1}. Note that two
metrics are used to evaluate the performance and a single low-pass filter is
used (Eq.\ref{eq:symetric-median}) and a threshold, i.e. $\tau = 0.11$. The end
of this section is devoted to a discussion about the choice of metric and its
relevance for the challenge.

% Datasets
To assess our method, we have used four datasets provided by the organizers
of the Connectomics challenge : \textit{normal-1, normal-2, normal-3, normal-4}.
Those datasets are time series of size $T=179500$ of image calcium from simulated
neuronal networks \citep{stetter2012model} with $|V|=1000$ neurons. The networks
to inferred have around 15000 edges and are similar to those
used to rank challenge participants.

% Metrics : ROC AUC and AUPRC
In supervised learning terms, the network inference task can be viewed as a
binary classification task where one has to correctly classify the presence
or the absence of edges of the graph. We assess the accuracy of the
network inference method using the area under the ROC curve (AUROC)
and the area under the precision-recall curve (AUPRC)
\citep{schrynemackers2013protocols}. Since neurons are not self-connected
($(i, i) \not \in E, \forall i \in V$) and some inference methods could
give maximal scores to the diagonal of $\hat{A}$, e.g. Pearson
correlation, we set those elements to the minimum of $\hat{A}$ prior
assessment.\\

On Tab.\ref{tab:tab3}, it can be seen that each step of our approach globally
improves the score on both metrics. However, processing, weights and averaging
(respectively denoted as $P$,$W$ and $Averaging$ in Tab.\ref{tab:tab3}) as
described in section \ref{sec:filter} are the key steps of our algorithm.

We compare our approach to the Pearson correlation (see Section
\ref{sec:inference}), the generalized transfer entropy (GTE) and GENIE3.

\textbf{GTE} \citep{stetter2012model}
, the baseline of the challenge, measures for all pairs of
neurons $i$ and $j$ the reduction of uncertainty in predicting future values of
$i$ given the observation of the time series of the neuron $j$
conditionally to the observation of past time series of neuron $i$,
whenever the average fluorescence is over a conditioning level.
Conditioning level was set to reproduce baseline of the challenge.

\textbf{GENIE3} \citep{huynhthu2010inferring} translates the inference of the
network of $p$ neurons into $p$ supervised learning tasks. For each task,
the fluorescence of a neuron is predicted from the fluorescence
of all other neurons using a tree based ensemble such as
random forest \citep{breiman2001random}. The tree ensemble associate a
variable importance scores \citep{louppe2013understanding}
to each neuron used in the predictions, which is used as a connectivity
score. The following parameters are used
$10000$ trees, $\log_2{p}$ neurons are randomly selected at each tree
node to split the data and the maximal tree depth is constrained
to 1, 2 or 3.\\

@@

%Opt: obviously when parameters are tuned/optimized, it's slightly better --> see appendix for a full description of the method
% Avg.: Averaging over the threholds and L-P filters improves a lot both metrics.

@@

\textcolor{red}{TODO : Discussion depth}

@@

@@

\begin{table}
\small
\centering
\begin{tabular}{@{}l *{8}{c}@{}}
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\cmidrule(r){2-5}\cmidrule(r){6-9}
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
\hline \hline
% Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
GENIE3 (depth = 1)     &  & & & & & & & \\
GENIE3 (depth = 2)     &  & & & & & & & \\
GENIE3 (depth = 3)     &  & & & & & & & \\
GTE                    &  & & & & & & & \\
% Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
\end{tabular}
\caption{TODO (network are undirected)}
\label{tab:tab2}
\end{table}

\begin{table}[tbh]
\centering
\caption{TODO}
\begin{tabular}{@{}l *{8}{c}@{}}
  & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
\cmidrule(r){2-5}\cmidrule(r){6-9}
Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
\midrule
No  filtering       & 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
P                   & 0.925 & 0.925 & 0.924 & 0.923 & 0.322 & 0.324 & 0.323 & 0.312\\
P + W               & 0.931 & 0.930 & 0.928 & 0.927 & 0.333 & 0.334 & 0.327 & 0.313\\
P + W + PCA         & 0.932 & 0.931 & 0.930 & 0.928 & 0.364 & 0.366 & 0.359 & 0.344\\
Averaging           & 0.943 & 0.942 & 0.942 & 0.939 & 0.403 & 0.404 & 0.398 & 0.388\\
Averaging + Directivity & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389\\
\end{tabular}
\end{table}

@@

\section*{Acknowledgements}

A. Joly and G. Louppe are research fellows of
the FNRS, Belgium.  A. Sutera is recipient of
a F.R.I.A. fellowship of F.R.S.-FNRS, Belgium.
This work is supported by PASCAL2 and the IUAP DYSCO, initiated by the
Belgian State, Science Policy Office.

>>>>

\section{Empirical experiments} \label{sec:results}

@@

% In this section, we will first describe datasets and methods used for the
% comparison. Then, on Table \ref{tab:tab3}, our approach is decomposed to show
% the contribution of each pre-processing steps. Afterwards, we will perform a comparison
% between several methods on Figure \ref{fig:curves} and Table \ref{tab:tab1}. Note that two
% metrics are used to evaluate the performance and a single low-pass filter is
% used (Eq. \ref{eq:symetric-median}) and a threshold, i.e. $\tau = 0.11$. The end
% of this section is devoted to a discussion about the choice of metric and its
% relevance for the challenge.

% \subsection{Assessment methodology}

% Datasets
To assess our method, we have used four datasets provided by the organizers
of the Connectomics challenge : \textit{normal-1, normal-2, normal-3, normal-4}.
Those datasets have $|V|=1000$ time series of length $T=179500$ of image
calcium from simulated neuronal networks \citep{stetter2012model}. Those
networks have approximately 15000 edges.

% Metrics : ROC AUC and AUPRC
In supervised learning terms, the network inference task can be viewed as a
binary classification task where one has to correctly classify the presence
or the absence of edges of the graph. We assess the accuracy of the
network inference method using the area under the ROC curve (AUROC)
and the area under the precision-recall curve (AUPRC)
\citep{schrynemackers2013protocols}. Since neurons are not self-connected
($(i, i) \not \in E, \forall i \in V$) and some inference methods could
give maximal scores to those edges, e.g. Pearson
correlation, those elements are set to have minimal connectivity score prior
assessment.

We will compare our approach to the Pearson correlation (see Section
\ref{sec:inference}), the generalized transfer entropy (GTE) and GENIE3:
\begin{itemize}

\item \textbf{GTE} \citep{stetter2012model}
, the baseline of the challenge, measures for all pairs of
neurons $i$ and $j$ the reduction of uncertainty in predicting future values of
$i$ given the observation of the time series of the neuron $j$
conditionally to the observation of past time series of neuron $i$,
whenever the average fluorescence is over a conditioning level.
Conditioning level was set to reproduce baseline of the challenge.

\item \textbf{GENIE3} \citep{huynhthu2010inferring} translates the inference of the
network of $p$ neurons into $p$ supervised learning tasks. For each task,
the fluorescence of a neuron is predicted from the fluorescence
of all other neurons using a tree based ensemble such as
random forest \citep{breiman2001random}. The tree ensemble associate a
variable importance scores \citep{louppe2013understanding}
to each neuron used in the predictions, which is used as a connectivity
score. The following parameters are used
$10000$ trees, $\log_2{p}$ variables are randomly selected to split a tree
node and the maximal tree depth is constrained to 1, 2 or 3.
\end{itemize}

\subsection*{Comparison with state-of-the art methods}

\begin{figure}[bth]
\centering
\includegraphics[width=0.6\linewidth]{images/curves}
\caption{Method precision-recall and roc curves.}
\label{fig:curves}
\end{figure}

@@

% Opt: obviously when parameters are tuned/optimized, it's slightly better -->
% see appendix for a full description of the method
% Avg.: Averaging over the threholds and L-P filters improves a lot both metrics.

@@

\textcolor{red}{TODO: depth discussion }

@@

% \begin{table}
% \small
% \centering
% \begin{tabular}{@{}l *{8}{c}@{}}
%   & \multicolumn{4}{c}{AUROC} & \multicolumn{4}{c}{AUPRC} \\
% \cmidrule(r){2-5}\cmidrule(r){6-9}
% Methods $\backslash$ normal- & 1 & 2 & 3 & 4 & 1 & 2 & 3 & 4 \\
% \hline \hline
% % Directivity \textcolor{red}{appendix?} & 0.535 & 0.526 & 0.533 & 0.535 & 0.012 & 0.012 & 0.012 & 0.012 \\
% GENIE3 (depth = 1)     &  & & & & & & & \\
% GENIE3 (depth = 2)     &  & & & & & & & \\
% GENIE3 (depth = 3)     &  & & & & & & & \\
% GTE                    &  & & & & & & & \\
% % Our approach + dir  \textcolor{red}{appendix?}   & 0.944 & 0.943 & 0.942 & 0.940 & 0.404 & 0.405 & 0.399 & 0.389 \\
% \end{tabular}
% \caption{TODO (network are undirected)}
% \label{tab:tab2}
% \end{table}

\subsection*{Signal processing highly improves performance}

On Table \ref{tab:tab3}, it can be seen that each step of our approach globally
improves the score on both metrics. However, processing, weights and averaging
(respectively denoted as $P$,$W$ and $Averaging$ in Tab.\ref{tab:tab3}) as
described in Section \ref{sec:filter} are the key steps of our algorithm.

@@

@@

\section*{Acknowledgements}
A. Joly and G. Louppe are research fellows of
the FNRS, Belgium.  A. Sutera is recipient of
a F.R.I.A. fellowship of F.R.S.-FNRS, Belgium.
This work is supported by PASCAL2 and the IUAP DYSCO, initiated by the
Belgian State, Science Policy Office.
